Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-17 paper_jasa.tex
proof-read to page 17

%\documentclass{article}
\documentclass[12pt]{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}

@@

\begin{tabular}{lp{5.5in}}
$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$. Covariates could include  demographic information of individuals, such as age, gender, education level etc. as well lineup-specific elements, e.g. effect size or difficulty level. \\
$\delta$ & is a vector of parameters for fixed effects.\\
$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
$\tau$ & is a vector of independent normally distributed random variables $\tau_{\ell i}$, with  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups. 
Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
\end{tabular}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. $n_c$ is the size of thecontaminated data. For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points. Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  
A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

Five replications of actual data sets for each slope value shown in Table \ref{tbl:experiment_params}, producing 30 lineups. Six groups of difficulty level based on slope are defined, difficulty = 0 for largest slope and difficulty = 5 for the smallest slope.    Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.

@@

%{\bf Screening scenarios:}
Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA. Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data. For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup. For the purposes of identifying these participants and cleaning the data, each block of 10 lineups that a subject received included several very easy ones. One of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly. The data of participants who did not answer correctly was removed. For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup. That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.

@@

Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.

\subsection{Model fitting}
For each parameter combination, the effect is computed as

\[
E=\beta (\sqrt {n}/\sigma).
\]

The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept. we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. Different subjects are expected to have different skill levels in evaluating lineups. To account for subject specific variation we use subject as random effect. We do not include random intercept, instead, we use an interaction between subject and effect $E$ to allow for subject specific random slope. This leads to setting  $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated in Equation \ref{eqn:mixed}.  While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power to be fixed at 0.05 for $E=0$ is not justified by the experiment 3 data.% (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!

For all the three experiments, each subject saw a lineup just once. This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. For computation we use R \citep{R} package lme4 \citep{lme4:2011}.

>>>>

%\documentclass{article}
\documentclass[12pt]{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}
\usepackage{hyperref}

@@

\begin{tabular}{lp{5.5in}}
$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$. Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g. effect size or difficulty level. \\
$\delta$ & is a vector of parameters for fixed effects.\\
$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
$\tau$ & is a vector of independent normally distributed random variables $\tau_{\ell i}$, with  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups. 
Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
\end{tabular}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. $n_c$ is the size of thecontaminated data. For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points. Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  
A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment  follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null  plots are scatterplots of null data generated by plotting simulated residuals from $N(0, {\hat{\sigma}}^2$)  against $X_1$.

Experiment three consists of a total of 30 lineups, made up of five replicates for each of the six slopes as shown in Table \ref{tbl:experiment_params}. We use the slope directly as a measure of difficulty, with difficulty = 0 for the largest slope and difficulty = 5 for the smallest slope.  Subjects were exposed to a total of ten lineups, with two lineups from each of the difficulty levels 0 through 3, and one  lineup each from levels 4 and 5.

@@

%{\bf Screening scenarios:}
Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA. Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data. For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup. For the purpose of identifying these participants and cleaning the data, we make use of one of the very easy lineups that every subject was exposed to.
For each subject we use a randomly selected easy lineup. If the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis. If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.

%
%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly. The data of participants who did not answer correctly was removed. 
%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup. That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.

@@

\subsection{Model fitting}
\noindent For each parameter combination, the effect is computed as

\[
E=\sqrt {n} \cdot \beta/\sigma.
\]

The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept., i.e. 
 $W_{\ell i} = E_{\ell i}$ in Equation \ref{eqn:mixed}. 
Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. Different subjects are expected to have different skill levels in evaluating lineups. We account for this by using subject-specific random  slopes for  effect ($E$).

For experiment 3 we so fit intercepts: both a fixed effect and subject-specific random intercepts, since forcing power
%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
to be fixed at 0.05 for $E=0$ is not justified by the data from experiment 3.% (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!

%For all the three experiments, each subject saw a lineup just once. This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.

The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. For computation we use R \citep{R} package lme4 \citep{lme4:2011}.
